Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex
mentioned baseline in proofs
Updated section 5

\section{Remarks on SVRPG}
The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $\alpha$ and $m$) that may be too conservative for practical applications.
Here we provide a practical and automatic way to choose the step size $\alpha$ and the number of sub-iterations $m$ performed between snapshots.
Additionally, we provide variants of SVRPG exploiting variance-reduction techniques for importance weights. Despite lacking theoretical guarantees, we will show that these methods can outperform the baseline SVRPG (Algorithm~\ref{alg:svrpg}).

\subsection{Zeroth Update}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\tilde{\vtheta}^s}$ feel like a waste of data. In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\tilde{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \tilde{\vtheta}^s + \alpha\gradApp{\tilde{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v_t^{s+1} \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the following, we will always assume this practical variant is employed.

\subsection{Meta-Parameter Selection}\label{sec:stopping}
% In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
% At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
The step size $\alpha$ is crucial to balance variance reduction and efficiency.
% The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
Epoch length $m$ influences the variance introduced by the importance weights. Low values of $m$ are associated with small variance but increase the frequency of snapshot points (which means many FG computations). High values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing large variance in the importance weights. We propose to adapt the two meta-parameters jointly.
% the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.

\textbf{Adaptive step size.}
A standard way to deal with noisy gradients is to use adaptive strategies to compute the step size.
ADAptive Moment estimation (ADAM)~\citep{kingma2014adam} stabilizes the parameter update by computing learning rates for each parameter based on an incremental estimate of the gradient variance.
Due to this feature, we would like to incorporate ADAM in the structure of the SVRPG update.
Recall that SVRPG performs two different updates of the parameters $\vtheta$: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
Given this structure, we suggest using two separate ADAM estimators:
\begin{align*}
        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} v_t^{s+1} \text{ for $t=1,\dots,m-1$},
\end{align*}
where $\alpha^{\textsc{FG}}_{s}$ is associated with the snapshot and $\alpha^{\textsc{SI}}_{s+1,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.
Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and w/o importance weights.
The use of two ADAM estimators allows to capture and exploit this property.

>>>>

\section{Remarks on SVRPG}
The convergence guarantees presented in the previous section come with requirements on the meta-parameters (\ie $\alpha$ and $m$) that may be too conservative for practical applications.
Here we provide a practical and automatic way to choose the step size $\alpha$ and the number of sub-iterations $m$ performed between snapshots.
Additionally, we provide a variant of SVRPG exploiting a variance-reduction technique for importance weights.
Despite lacking theoretical guarantees, we will show in Section~\ref{sec:exp} that this method can outperform the baseline SVRPG (Algorithm~\ref{alg:svrpg}).

\subsection{Full Gradient Update}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\wt{\vtheta}^s}$ feels like a waste of data (the term $\sum_i g(\tau_i) - \omega(\tau_i) g(\tau_i) =0$ since $\vtheta_0 = \wt{\vtheta}$).
In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\wt{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \wt{\vtheta}^s + \alpha\gradApp{\wt{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha 
%        v_t^{s+1}
        \blacktriangledown J(\vtheta^{s+1}_t)
        \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the following, we will always assume this practical variant is employed.

\subsection{Meta-Parameter Selection}\label{sec:stopping}
% In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
% At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
The step size $\alpha$ is crucial to balance variance reduction and efficiency, while
% The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
the epoch length $m$ influences the variance introduced by the importance weights. Low values of $m$ are associated with small variance but increase the frequency of snapshot points (which means many FG computations). High values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing large variance in the importance weights. We propose to adapt the two meta-parameters jointly.
% the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.

\textbf{Adaptive step size.}
A standard way to deal with noisy gradients is to use adaptive strategies to compute the step size.
ADAptive Moment estimation (ADAM)~\citep{kingma2014adam} stabilizes the parameter update by computing learning rates for each parameter based on an incremental estimate of the gradient variance.
Due to this feature, we would like to incorporate ADAM in the structure of the SVRPG update.
Recall that SVRPG performs two different updates of the parameters $\vtheta$: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
Given this structure, we suggest using two separate ADAM estimators:
\begin{align*}
        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} 
%        v_t^{s+1}
        \blacktriangledown J(\vtheta^{s+1}_t)
        \text{ for $t=1,\dots,m-1$},
\end{align*}
where $\alpha^{\textsc{FG}}_{s}$ is associated with the snapshot and $\alpha^{\textsc{SI}}_{s+1,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.
Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and w/o importance weights.
The use of two ADAM estimators allows to capture and exploit this property.
